Evaluating infectious disease forecasts \\ with allocation scoring rules

### 2024-02-16 alloscore_manuscript/alloscore-application-shorter.Rnw
more edits and reformatting... Discussion paragraphs 2,3,5 still need work.

Despite this now well-developed discussion of the quality-value distinction in the larger forecasting community, we are
aware of only a limited literature attempting to connect the value of \emph{infectious disease} forecasts to their
impact on and through policy.  And within this body of work we have found the discussion of such a connection to still
be at a formally and quantitively imprecise stage. In \cite{ioannidis2022forecastingCOVIDfailed}, the possible negative
consequences of inaccurate forecasts of infectious disease are discussed, but there is no attempt to quantify the
utility or loss incurred as a result of those forecasts. \cite{bilinski_adaptive_2023} explore ways in which predictive
classifiers of local COVID-19 risk levels in the US could be tuned to policymaker preferences for different costs
associated with over- and under-reaction to disease dynamics, but they do not clearly identify the source of these
costs or how they depend on quantifiable policy choices. A similar discussion related to dengue countermeasures in
Vietnam appears in \cite{colon-gonzalez_probabilistic_2021}.  A novel version of the WIS informally motivated by
utility considerations is developed in \cite{marshall2023predictions}, but the score is not derived in a
decision-theoretic manner.  There is also a thread of literature that frames infectious disease modeling as a component
of a larger system for understanding how policy goals, means, and choices interact and constrain one another. As an
example, \cite{Probert2016decisionMakingFootMouth} explore how policy recommendations ought to flow from a possibly 
incongruous set of simulation-based projection models of a hypothetical foot-and-mouth disease outbreak when there are 
ranges of plausible responses and stakeholder interests. Decision theory plays a prominent role here, but not explicitly
as a way to evaluate the choices made in developing the models.

@@

Decision theory is concerned with understanding how an \emph{action} $x$ both is and ought to be selected from a set of
possible actions $\mathcal{X}$ --- that is, how a decision is and ought to be made --- based on the anticipated
consequences of $x$ under future states. We represent such future states as a random variable $Y$ and refer to a
realization $y$ of $Y$ as an \emph{outcome}. In this paper $x$ will be related to the level of investment in a measure
designed to mitigate severe disease outcomes such as hospital beds, ventilators, medication, or medical staff, with
$\mathcal{X}$ being the set of all feasible levels of investment, while an outcome $y$ will correspond to the number of
individuals who eventually become sick and would benefit from the mitigation measure. A basic strategy in decision
theory is to assume that the consequences of an action $x$ under outcome $y$ can be assigned a numerical value, or
\emph{loss}, measuring the success of $x$, which in our examples will be the extent to which $x$ meets the realized need
$y$.  One way that a decision maker can try to anticipate the consequence of an action $x$ is by using a forecast $F$ of
$Y$. The loss of the action $x$ chosen according to $F$ can, in turn, be intepreted as the value of $F$ as an input to
the decision making process under the realized outcome $y$ of $Y$.

We arrive through this discussion at a three-step procedure for developing scoring rules for probabilistic forecasts:
\begin{enumerate}
  \item Specify a \emph{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$
    eventually occurs.
  \item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under
    the distribution $F$.
  \item The \emph{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: 
    $S(F, y) = s(x^F, y)$.
\end{enumerate}
This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss
function. We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} and in section
\ref{sec:proper} of the supplement we review the basic fact from the scoring rule literature (e.g.,
\cite{dawid2007geometry,gneiting2007strictly}) that Bayes scoring rules are proper by construction.

@@

It is helpful to clearly distinguish between the time $t_d$ when a \emph{d}ecision is made about a public health
resource allocation and the time $t_r$ when \emph{r}esource needs that might be addressed by that allocation occur. Our
setup assumes that $t_d < t_r$ and that the resource in question can only meet the need realized at $t_r$, not prevent
it. That is, we stipulate that the random vector $Y$ of need is independent from the chosen allocation $x$. In the
context of infectious disease, this means that we do not consider resources, such as vaccines, that are intended to
reduce the number of people who will become sick at some point in the future. Instead, our setup addresses resources
like hospital beds, oxygen supply, or ventilators which are intended to meet the medical needs of patients who are
already sick. Additionally, our setup addresses decision-making that is related to resource needs only at the time
$t_r$; we do not consider sequences of multiple decisions that are made over time or account for the impact of decisions
on resource needs at any time other than $t_r$. We outline some opportunities to extend our work to more complex
decision making settings in the discussion.

@@

We illustrate with an application to hospital admissions in the U.S., considering a hypothetical problem of allocating a
limited supply of medical resources to states.

@@

%\subsubsection{Hospitalization data}
%Starting in the summer of 2020, the US Health and Human Services (HHS) began reporting counts of daily new admissions to hospitals for individuals with COVID-19 \citep{healthdatagov_covid-19_nodate}.
%These daily counts were available for the US as a whole, and all states and several additional jurisdictions such as Puerto Rico and Washington DC.
%The data were updated daily and were available for download by the public through the HHS HealthData.gov website.
%For this analysis, we downloaded the hospitalizations data through the covidHubUtils R package, which connects users to the most recent version of the data \cite{wang-covidhubutils}.

%\subsubsection{Forecast data}
The US COVID-19 Forecast Hub collected short-term forecasts of daily new hospital admissions for individuals with COVID-19 starting in December 2020 \citep{cramer_united_2022}.
The target data for these forecasts were hospital admissions as reported by the US Department of Health and Human Services through the HealthData.gov website.
Forecasts were probabilistic predictions of the number of new hospital admissions on a particular day in the future, in a specific jurisdiction of the US (national level, state, or territory).
Probability distributions were represented using a set of 23 quantiles for each individual prediction, including a median and the lower and upper limits of 11 central prediction intervals, from a 99\% to a 10\% prediction interval.

The analysis in this work focuses on forecasts made before and during the first wave of the Omicron SARS-CoV-2 variant in the US.
As such, we analyzed forecasts for the 15 weeks starting with Monday November 22, 2021 through Monday February 28, 2022.
%Forecast data were downloaded on for analysis on \Sexpr{format(as.Date(metadata$time[which(metadata$name == "forecast_data")]), "%B %d, %Y")}.

@@

We measured forecast skill using two forecast scores, the allocation score (AS) and the weighted interval score (WIS), both defined above.
We computed these scores for the 14 day ahead forecasts made each week.
For this analysis, we fixed the resource constraint used for the AS to be $K=15,000$, based roughly on a reported number of ventilators available for reallocation in the US \citep{ajao_assessing_2015}.
We computed the mean WIS (MWIS) across all of the forecasted state-level locations.
%For each week, we computed the allocation score for the 14 day-ahead forecast.

For both scores, we also computed models standardized rank among all models that submitted forecasts each week. The standardized rank is between 0 and 1, where 0 corresponds to the worst rank and 1 to the best. In the case of a tie between one or more models, all models received the better rank.

As described above, predictions were submitted to the Forecast Hub in the form of a set of 23 quantiles of the predictive distribution. The WIS can be directly calculated from these quantiles. However, calculation of the AS requires that the full cumulative distribution functions of the forecast distributions for each location are available. For the purpose of this analysis, to calculate the allocation score we approximated the full cumulative distribution functions based on the provided quantiles. On the interior of the provided quantiles we used a monotonic cubic spline to interpolate the quantiles, and in the lower and upper tails we used normal distributions with parameters selected so as to match the two lowest and two highest quantiles (see section \ref{sec:distfromq} the supplement for further details). In 
section \ref{sec:propriety_of_parametric_approximation} of the supplement, we show that if this evaluation procedure had been specified prospectively, the resulting score would be proper \textemdash but a post hoc application of this procedure is improper. We use the procedure here to illustrate the properties of the AS, and note that a collaborative forecasting hub interested in using the AS for evaluation could circumvent issues with propriety by communicating the definition of the AS to forecasters and then collecting allocations at specified resource levels as part of forecast submissions.

@@

We used two benchmark methods as references to allow for comparison of model-derived allocations with existing standard practices.
First, we evaluated forecasts from the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk \citep{cramer_evaluation_2022}.
Second, we generated a proposed set of allocations where the quantity allocated to each state was proportional to that state's population (using US Census data of vintage 2022 \citep{us_census_nst_est2022}), referred to below as \texttt{per-capita}.
These two approaches generally reflect common choices for ``best practices'' of allocation: either allocating resources based on the most recent observed data, or in proportion to the population of each location.
%As such, they serve as good comparisons for allocations based on forecast models.

@@

model_palette <- setNames(
  hcl(h = seq(15, 375, length = 18), l = 65, c = 100),
  c(unique(forecast_data$model), "per-capita"))
model_palette["JHUAPL-Bucky"] <- "thistle"
model_palette["CU-select"] <- "wheat"

@@

@@

To illustrate the mechanics of allocation scoring, we start by focusing on how forecasts generated on or before December 20, 2021, with predictions for January 03, 2022, were scored by different metrics.
This week was around the peak of the Omicon wave nationally, with individual states typically observing a peak at or after January 3, 2022.

%For many locations, forecasts predicted lower values than were eventually observed, as this was during the period of steep increase of viral transmission across many states.
Of the 10 models evaluated for this one week, the CU-select model had the most accurate forecasts according to the allocation score while the USC-SI\_kJalpha model had the most accurate forecasts based on MWIS (Table \ref{tab:multi-k-scores}).
The JHUAPL-Bucky model had the second best MWIS but the third worst allocation score.

@@

left_join(ias, as) |>
  left_join(wis_vals, by="model") |>
  rename(Model = model,
         `IAS centered at 15k` = IAS_wt,
         `IAS uniform` = IAS) |>
  select(Model, AS, MWIS, `IAS centered at 15k`, `IAS uniform`) |>
  arrange(AS) |>
  xtable::xtable(caption = "For one illustrative week, a comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS). All metrics are shown for 10 models that made forecasts of hospital admissions for 2022-01-03. Results are sorted by AS. For all metrics, lower scores indicate better accuracy.",
                 digits = 0,
                 label = "tab:multi-k-scores") |>
  print(include.rownames = FALSE, NA.string = "-")
@

@@

%Forecasts and allocations for a selection of states with high numbers of hospitalizations on January 3, 2022 reveal mechanics about how allocations are made (Figure \ref{fig:thermometer-plot}).
%As described in the methods above, allocations for a given location from a specific model are generated by finding the single quantile of the probabilistic forecast across all locations that reaches the resource allocation limit (in this case, assumed to be 15,000).
%A direct comparison between the forecasts from the JHUAPL-Bucky and CU-select models show that while the JHUAPL-Bucky forecast distributions were closer to the eventual observations in many states, the allocations suggested by those forecasts often were more inefficient than those from the CU-select model (Figure \ref{fig:thermometer-plot}).

@@

%Together, these results corroborate the understanding that the allocation score and weighted interval score measure different aspects of forecast performance. The allocation score penalizes models that mischaracterize the relative magnitude of resource need across different locations, while WIS penalizes forecasts that do not capture the absolute magnitude of the target quantity in each location.

@@

<<metrics-over-time, fig.height=8, fig.width=8, fig.cap="Hospital admissions and evaluation metrics over time. Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022. These are the dates for which forecasts were made and evaluated. A horizontal dashed line at 15,000 shows the hypothetical resource constraint $K$. Panel B shows allocation scores (AS) for each model's 14 day-ahead forecast, across all US states. The $x$-axis corresponds to the date of the observation that a model's prediction was targeting (e.g., the date the forecast was made plus the forecast horizon). AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down). In Panel B, the per-capita allocation is drawn in a heavier solid line. Panel C shows the MWIS metric across weeks, averaged across all states. MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the peak of the Omicron wave.">>=
p2 <- ggplot(alloscore_data_summaries,
             aes(x=target_end_date, y=alloscore, color=model, shape=model)) +
  scale_x_date(limits=range(score_data_summaries$target_end_date)) +
  scale_color_manual(values = model_palette) +
  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4, 6)) +
  scale_linewidth_manual(values = c(.5,1.5)) +
  ylab("Allocation score") +
  geom_point() + geom_line(aes(linewidth = as.factor(model == "per-capita")))+
  theme(legend.pos="none") +
  xlab(NULL)

p4 <- ggplot(bind_rows(score_data_summaries, percap %>%
                         select(model, reference_date) %>%
                         mutate(reference_date = as.Date(reference_date))),
             aes(x=target_end_date, y=mwis, color=model,shape=model)) +
  geom_point() + geom_line() +
  scale_color_manual(values = model_palette) +
  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4, 6)) +
  xlab("Forecast target date") +
  ylab("MWIS")+
  theme(legend.pos="bottom")+
  guides(colour = guide_legend(title=NULL, ncol = 4),
         shape = guide_legend(title=NULL, ncol = 4))

@@

Allocation scores varied substantially by date and by model (Figure \ref{fig:metrics-over-time}).
For predictions made for the first three Mondays in December 2021 and the last three Mondays in February 2022 all models had allocation scores under 500 (and the mean across all models was less than 100), indicating that unnecessary unmet need was fairly low on those days.
The allocation scores were on the whole highest when the observed number of new hospital admissions was closest to the resource threshold of 15,000, as those are the times when any mistakes in allocation are costly in terms of wasting resources in one location that could have been used in another.
%Predictions made during the peak week and just after showed the highest variation in allocation scores, with some models having allocation scores under 1000 and others having values over 3500.

@@

Most models were ranked similarly when evaluated using mean allocation scores (MAS) and mean weighted interval scores (MWIS), with some places of disagreement.
The four most accurate and the five least accurate models were the same according to both metrics, although not in exactly the same order (Table \ref{tab:multi-week-performance-summary}).
(This comparison excludes the per-capita allocation which does not use forecasts and therefore cannot have a WIS.)
Notably, the best model according to MWIS was the JHUAPL-SLPHospEns model, but this model had only the fourth best MAS.

<<table-score-summary-all-weeks, results='asis'>>=
all_scores |>
  filter( !(reference_date.x %in% c("2021-11-22", "2022-02-28"))) |>
  group_by(model) |>
  summarize(N = sum(!is.na(alloscore)),
            MAS = mean(alloscore, na.rm=TRUE),
            MWIS = mean(mwis, na.rm=TRUE)) |>
  filter(N > 12) |>
  arrange(MAS) |>
  select(model, MAS, MWIS) |>
  xtable::xtable(caption = "Mean weighted interval scores (MWIS) and mean allocation scores (MAS) by model across 13 weeks. These results show the average performance across time for the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, as well
  as for a per-capita allocation. WIS is not defined for the per-capita allocation which does not use forecasts.",
                 digits = 0,
                 label = "tab:multi-week-performance-summary") |>
  print(include.rownames = FALSE, NA.string = "-")
@

@@

@@

Models showed differing levels of correlation between their allocation scores and MWIS values (Figure \ref{fig:metrics-correlation}), explaining in part why aggregate MWIS and MAS values did not always agree.
Here are some examples of the different model-specific patterns observed:
\begin{itemize}
\item Several models showed a positive association between allocation score and MWIS ranks (e.g., \texttt{Karlen-pypm} and \texttt{USC-SI\_kJalpha}).
\item One model had consistently strong MWIS ranks but also had highly variable allocation score ranks with no clear association between the two (e.g., \texttt{JHUAPL-SLPHospEns})
\item One model performed consistently well for both metrics (\texttt{COVIDhub-ensemble})
\item One model performed consistently well for allocation score but had only middling ranks for MWIS (\texttt{CMU-TimeSeries})
\end{itemize}

@@

\subsubsection{Integrated allocation score across values of K}

@@

AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of hospitalizations on January 3, 2021 as well as the per-capita allocation (Figure \ref{fig:multi-k}A).
These calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of \Sexpr{format(tot_y, big.mark = ",")} that day.
Rankings of AS from all models were fairly stable across a range of values for K, with some crossings, especially at $K$ values further away from the observed value.

The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the constraint ($K$), possibly taking weights into account for different values of $K$ that might be more or less likely (Section \ref{sec:methods.detailed.integrated_allocation}).
IAS was computed for two distributions on $K$, one uniform across the entire range and the other a symmetric distribution around 15,000 (Figure \ref{fig:multi-k}B).
Both versions of the IAS were correlated with the original AS conducted at $K=15,000$, with the higher correlation coming from the distribution that was centered at $K=15,000$ (Figure \ref{fig:multi-k}C).
Model rankings based on the AC and the centered IAS were roughly similar, with the top and bottom three approaches being the same for both scores (Table \ref{tab:multi-k-scores}).

<<multi-k, fig.height=5, fig.width = 8, fig.cap = "Allocation scores (AS) across different resource constraints ($K$) for 10 models that made forecasts on 2021-12-20. Panel A shows, for each model, the AS for values of $K$ between 200 and 60,000 at increments of 200. The AS show a sharp peak just under 20,000, near the eventually observed number of hospitalizations. Two possible weighting functions for the Integrated Allocation Score (IAS) are shown. The first (blue shaded area) computes weights proportional to a normal distribution centered at 15,000 with a standard deviation of 3,000, and truncated to be between 5,000 and 25,000. The second (orange shaded area) uses a uniform weight for all possible values of $K$. Note that the AS used in earlier sections of the application uses the single fixed value of $K=15,000$.  Panel B shows how the two versions of the IAS ($y$-axis) correlate with the AS at $K=15,000$ ($x$-axis). Every point represents the AS at 15,000 and a version of the IAS for one model. As we would expect, the IAS centered at 15,000 (circles) is more closely correlated with the AS at 15,000 than is the uniform IAS (triangles).">>=

>>>>

Despite this now well-developed discussion of the quality-value distinction in the larger forecasting community, we are
aware of only a limited literature attempting to connect the value of \emph{infectious disease} forecasts to their
impact on and through policy.  And within this body of work we have found the discussion of such a connection to still
be at a formally and quantitatively imprecise stage. In \cite{ioannidis2022forecastingCOVIDfailed}, the possible negative
consequences of inaccurate forecasts of infectious disease are discussed, but there is no attempt to quantify the
utility or loss incurred as a result of those forecasts. \cite{bilinski_adaptive_2023} explore ways in which predictive
classifiers of local COVID-19 risk levels in the US could be tuned to policymaker preferences for different costs
associated with over- and under-reaction to disease dynamics, but they do not clearly identify the source of these
costs or how they depend on quantifiable policy choices. A similar discussion related to dengue countermeasures in
Vietnam appears in \cite{colon-gonzalez_probabilistic_2021}.  A novel version of the WIS informally motivated by
utility considerations is developed in \cite{marshall2023predictions}, but the score is not derived in a
decision-theoretic manner.  There is also a thread of literature that frames infectious disease modeling as a component
of a larger system for understanding how policy goals, means, and choices interact and constrain one another. As an
example, \cite{Probert2016decisionMakingFootMouth} explore how policy recommendations ought to flow from a possibly 
incongruous set of simulation-based projection models of a hypothetical foot-and-mouth disease outbreak when there are 
ranges of plausible responses and stakeholder interests. Decision theory plays a prominent role here, but not explicitly
as a way to evaluate the choices made in developing the models.

@@

Decision theory is concerned with understanding how an \emph{action} $x$ both is and ought to be selected from a set of
possible actions $\mathcal{X}$ --- that is, how a decision is and ought to be made --- based on the anticipated
consequences of $x$ under future states. We represent such future states as a random variable $Y$ and refer to a
realization $y$ of $Y$ as an \emph{outcome}. In this paper $x$ will be related to the level of investment in a measure
designed to mitigate severe disease outcomes such as hospital beds, ventilators, medication, or medical staff, with
$\mathcal{X}$ being the set of all feasible levels of investment, while an outcome $y$ will correspond to the number of
individuals who eventually become sick and would benefit from the mitigation measure. A basic strategy in decision
theory is to assume that the consequences of an action $x$ under outcome $y$ can be assigned a numerical value, or
\emph{loss}, measuring the success of $x$, which in our examples will be the extent to which $x$ meets the realized need
$y$.  One way that a decision maker can try to anticipate the consequence of an action $x$ is by using a forecast $F$ of
$Y$. The loss of the action $x$ chosen according to $F$ can, in turn, be interpreted as the value of $F$ as an input to
the decision making process under the realized outcome $y$ of $Y$.

We arrive through this discussion at a three-step procedure for developing scoring rules for probabilistic forecasts:
\begin{enumerate}
  \item Specify a \emph{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$
    eventually occurs.
  \item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under
    the distribution $F$.
  \item The \emph{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: 
    $S(F, y) = s(x^F, y)$.
\end{enumerate}
This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss
function. We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} --- a terminology we find to
be convenient but which is not, to our knowledge, standard.  In section \ref{sec:proper} of the supplement we review the
basic fact from the scoring rule literature (e.g., \cite{dawid2007geometry,gneiting2007strictly}) that Bayes scoring
rules are proper by construction.

@@

It is helpful to clearly distinguish between the time $t_d$ when a \emph{d}ecision is made about a public health
resource allocation and the time $t_r$ when \emph{r}esource needs that might be addressed by that allocation occur. Our
setup assumes that $t_d < t_r$ and that the resource in question can only meet the need realized at $t_r$, not prevent
it. That is, we stipulate that the random vector $Y$ of need is independent of the chosen allocation $x$. In the
context of infectious disease, this means that we do not consider resources, such as vaccines, that are intended to
reduce the number of people who will become sick at some point in the future. Instead, our setup addresses resources
like hospital beds, oxygen supply, or ventilators which are intended to meet the medical needs of patients who are
already sick. Additionally, our setup addresses decision-making that is related to resource needs only at the time
$t_r$; we do not consider sequences of multiple decisions that are made over time or account for the impact of decisions
on resource needs at any time other than $t_r$. We outline some opportunities to extend our work to more complex
decision making settings in the discussion.

@@

We illustrate our new forecast evaluation framework with an application to hospital admissions in the U.S., considering
a hypothetical problem of allocating a limited supply of medical resources to states.

@@

The US COVID-19 Forecast Hub collected short-term forecasts of daily new hospital admissions for individuals with
COVID-19 starting in December 2020 \citep{cramer_united_2022}. The target data for these forecasts were hospital
admissions as reported by the US Department of Health and Human Services through the HealthData.gov website. Forecasts
were probabilistic predictions of the number of new hospital admissions on a particular day in the future, in a specific
jurisdiction of the US (national level, state, or territory). Probability distributions were represented using a set of
23 quantiles for each individual prediction, including a median and the lower and upper limits of 11 central prediction
intervals, from a 99\% to a 10\% prediction interval.

The analysis in this work focuses on forecasts made before and during the first wave of the Omicron SARS-CoV-2 variant
in the US. As such, we analyzed forecasts for the 15 weeks starting with Monday November 22, 2021 through Monday
February 28, 2022.

@@

We evaluated forecasts using the allocation score (AS) and the weighted interval score (WIS), computed at a horizon of
14 days. For this analysis, we fixed the resource constraint used for the AS to be $K=15,000$, based roughly on a
reported number of ventilators available for reallocation in the US \citep{ajao_assessing_2015}. We computed the mean
WIS (MWIS) across all of the forecasted state-level locations.

For both scores, we also computed standardized ranks among all models that submitted forecasts each week. The
standardized rank lies in the interval $[0, 1]$, where 0 corresponds to the worst rank and 1 to the best. In the case of
a tie between one or more models, all models received the better rank.

As described above, predictions were submitted to the Forecast Hub in the form of a set of 23 quantiles of the
predictive distribution. The WIS can be directly calculated from these quantiles. However, calculation of the AS
requires that the full cumulative distribution functions of the forecast distributions for each location are available.
For the purpose of this analysis, to calculate the allocation score we approximated the full cumulative distribution
functions based on the provided quantiles. On the interior of the provided quantiles we used a monotonic cubic spline to
interpolate the quantiles, and in the lower and upper tails we used normal distributions with parameters selected so as
to match the two lowest and two highest quantiles (see section \ref{sec:distfromq} the supplement for further details).
In section \ref{sec:propriety_of_parametric_approximation} of the supplement, we show that if this evaluation procedure
had been specified prospectively, the resulting score would be proper, \textemdash but that a post hoc application of this
procedure is improper. We use the procedure here to illustrate the properties of the AS, and note that a collaborative
forecasting hub interested in using the AS for evaluation could circumvent propriety issues by communicating the
definition of the AS to forecasters and then collecting allocations at specified resource levels as part of forecast
submissions.

@@

In order to explore how model-derived allocations might be compare to ``standard operating procedures'' for allocation
used by public health officials, we scored two simple benchmark methods. First, we evaluated forecasts from the
\texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds
based on a random walk \citep{cramer_evaluation_2022}. Second, we generated a proposed set of allocations where the
quantity allocated to each state was proportional to that state's population (using US Census data of vintage 2022
\citep{us_census_nst_est2022}), referred to below as \texttt{per-capita}. These two approaches generally reflect common
choices for ``best practices'' of allocation: either allocating resources based on the most recent observed data, or in
proportion to the population of each location.

@@

@@

<<normalize-model-aesthetics>>=
model_palette <- setNames(
  hcl(h = seq(15, 375, length = 18), l = 65, c = 100),
  c(unique(forecast_data$model), "per-capita"))

model_palette["JHUAPL-Bucky"] <- "thistle"
model_palette["CU-select"] <- "wheat"

model_linewidths <- setNames(rep(.5,
  length.out = length(model_palette)
), names(model_palette))
model_linewidths["per-capita"] <- 1.5
@

@@

To illustrate the mechanics of allocation scoring, we start by focusing on how forecasts generated on or before December
20, 2021, with predictions for January 03, 2022, were scored by different metrics. This week was around the peak of the
Omicron wave nationally, with individual states typically observing a peak at or after January 3, 2022.

Of the 10 models evaluated for this one week, the CU-select model had the most accurate forecasts according to the
allocation score while the USC-SI\_kJalpha model had the most accurate forecasts based on MWIS (Table
\ref{tab:multi-k-scores}). The JHUAPL-Bucky model had the second best MWIS but the third worst allocation score.

@@

left_join(ias, as) |>
  left_join(wis_vals, by="model") |>
  rename(Model = model,
         `IAS centered at 15k` = IAS_wt,
         `IAS uniform` = IAS) |>
  select(Model, AS, MWIS, `IAS centered at 15k`, `IAS uniform`) |>
  arrange(AS) |>
  xtable::xtable(caption = paste("For one illustrative week, a comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS), (see Section \\ref{sec:ias_examp}). All metrics are shown for 10 models that made forecasts of hospital admissions for 2022-01-03. Results are sorted by AS. For all metrics, lower scores indicate better accuracy."),
                 digits = 0,
                 label = "tab:multi-k-scores") |>
  print(include.rownames = FALSE, NA.string = "-")
@

@@

@@

@@

<<metrics-over-time, fig.height=8, fig.width=8, fig.cap="Hospital admissions and evaluation metrics over time. Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022. These are the dates for which forecasts were made and evaluated. A horizontal dashed line at 15,000 shows the hypothetical resource constraint $K$. Panel B shows allocation scores (AS) for each model's 14 day-ahead forecast, across all US states. The $x$-axis corresponds to the date of the observation that a model's prediction was targeting (e.g., the date the forecast was made plus the forecast horizon). AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down). In Panel B, the per-capita allocation is drawn in a heavier solid line. Panel C shows the MWIS metric across weeks, averaged across all states. MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the peak of the Omicron wave.">>=
p2 <- ggplot(alloscore_data_summaries,
             aes(x=target_end_date, y=alloscore, color=model, shape=model, linewidth = model)) +
  geom_point() + geom_line()+
  scale_x_date(limits=range(score_data_summaries$target_end_date)) +
  scale_color_manual(values = model_palette) +
  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4, 6)) +
  scale_linewidth_manual(values = model_linewidths) +
  ylab("Allocation score") +
  theme(legend.pos="none") +
  xlab(NULL)

p4 <- ggplot(bind_rows(score_data_summaries, percap %>%
                         select(model, reference_date) %>%
                         mutate(reference_date = as.Date(reference_date))),
             aes(x=target_end_date, y=mwis, color=model,shape=model, linewidth = model)) +
  geom_point() + geom_line() +
  scale_color_manual(values = model_palette) +
  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4, 6)) +
  scale_linewidth_manual(values = model_linewidths) +
  xlab("Forecast target date") +
  ylab("MWIS")+
  guides(colour = guide_legend(title=NULL, ncol = 4),
         shape = guide_legend(title=NULL, ncol = 4),
         linewidth = guide_legend(title=NULL, ncol = 4)) +
  theme(legend.pos="bottom")

@@

Allocation scores varied substantially by date and by model (Figure \ref{fig:metrics-over-time}). For predictions made
for the first three Mondays in December 2021 and the last three Mondays in February 2022 all models had allocation
scores under 500 (and the mean across all models was less than 100), indicating that unnecessary unmet need was fairly
low on those days. The allocation scores were on the whole highest when the observed number of new hospital admissions
was closest to the resource threshold of 15,000.  It was on these occasions that models were most likely to waste resources
in one location that the oracle would have put to use in another.

@@

Model rankings on mean allocation score (MAS) and time-averaged mean weighted interval score
(taMWIS) were broadly similar, with some places of disagreement. The four most accurate and the five least accurate models were the same
according to both metrics, although not in exactly the same order (Table \ref{tab:multi-week-performance-summary}).
(This comparison excludes the per-capita allocation which does not use forecasts and therefore cannot be assigned a MWIS.)
Notably, the JHUAPL-SLPHospEns model ranked best on taMWIS but had only the fourth best MAS.

<<table-score-summary-all-weeks, results='asis'>>=
all_scores |>
  filter( !(reference_date.x %in% c("2021-11-22", "2022-02-28"))) |>
  group_by(model) |>
  summarize(N = sum(!is.na(alloscore)),
            MAS = mean(alloscore, na.rm=TRUE),
            taMWIS = mean(mwis, na.rm=TRUE)) |>
  filter(N > 12) |>
  mutate(`taMWIS rank` = rank(taMWIS, na.last = "keep")) |>
  arrange(MAS) |>
  select(model, MAS, taMWIS, `taMWIS rank`) |>
  xtable::xtable(caption = "Time averaged mean weighted interval scores (taMWIS) and mean allocation scores (MAS) by model across 13 weeks. These results show the average performance across time for the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, as well
  as for a per-capita allocation. MWIS is not defined for the per-capita allocation which does not use forecasts.",
                 digits = 0,
                 label = "tab:multi-week-performance-summary") |>
  print(include.rownames = FALSE, NA.string = "-")
@

@@

We can gain some insight into the discordance between MAS and taMWIS by examining the correlation between time-specific AS and MWIS values of the various models in Figure
\ref{fig:metrics-correlation}. We find, for example, no clear association between the consistent
and high MWIS rank of the JHUAPL-SLPHospEns model and its highly variable AS rank. This contrasts 
with a clearly positive correlation between MWIS and AS ranks in other models such as 
Karlen-pypm and USC-SI\_kJalpha.
We also can see in more detail the consistently high performance COVIDhub-ensemble on both scores.
And while it did not submit forecasts for enough weeks to be included in Table \ref{tab:multi-week-performance-summary}, CMU-TimeSeries is an interesting example of a model that 
performed consistently well on AS but had only middling MWIS ranks.

@@

@@

\subsubsection{Integrated allocation score across values of K}
\label{sec:ias_examp}

@@

AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of
hospitalizations on January 3, 2021 as well as for the per-capita allocation (Figure \ref{fig:multi-k}A). These
calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of
\Sexpr{format(tot_y, big.mark = ",")} that day. Model ranks by AS were fairly stable across all $K$, and there appears
to be an substantial interval around the observed value over which ranks are constant.

The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the
constraint ($K$), possibly assigning higher weights to more likely values of $K$ (Section
\ref{sec:methods.detailed.integrated_allocation}). IAS was computed for two weight distributions on $K$, one uniform
across the entire range and the other a symmetric distribution around 15,000 (the orange and blue shaded areas in Figure
\ref{fig:multi-k}A, respectively). Both versions of the IAS were correlated with the original AS for $K=15,000$, with
the higher correlation coming from the weighting centered at $K=15,000$ (Figure \ref{fig:multi-k}B). Model rankings
based on the AS and the centered IAS were roughly similar, with the top and bottom three approaches being the same for
both scores (Table \ref{tab:multi-k-scores}).

<<multi-k, fig.height=4, fig.width = 8, fig.cap = "Allocation scores (AS) across different resource constraints ($K$) for 10 models that made forecasts on 2021-12-20. Panel A shows, for each model, the AS for values of $K$ between 200 and 60,000 at increments of 200. The AS show a sharp peak just under 20,000, near the eventually observed number of hospitalizations. Two possible weighting functions for the Integrated Allocation Score (IAS) are shown. The first (blue shaded area) computes weights proportional to a normal distribution centered at 15,000 with a standard deviation of 3,000, and truncated to be between 5,000 and 25,000. The second (orange shaded area) uses a uniform weight for all possible values of $K$. Note that the AS used in earlier sections of the application uses the single fixed value of $K=15,000$.  Panel B shows how the two versions of the IAS ($y$-axis) correlate with the AS at $K=15,000$ ($x$-axis). Every point represents the AS at 15,000 and a version of the IAS for one model. As we would expect, the IAS centered at 15,000 (circles) is more closely correlated with the AS at 15,000 than is the uniform IAS (triangles).">>=
